---
title: "Assignment 4"
author: Maimoona Mohsin Subhani 
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r}
library (tidyverse)
energy <- read_csv("data/owid_energy.csv") 

```
Focusing on renewable energy in seven countries from 2000 onward 
```{r}
energy_sub <-energy |>
    filter(country %in% c("United States", "China", "India", "United Kingdom", "Norway", "Pakistan", "Ghana"),
         year >= 2000)
```

### Graph 1: Renewable share of electricity

```{r}
energy_sub |>
  ggplot(aes(x = year, y = renewables_electricity_share_pct, color = country)) +
  geom_line(linewidth = 1, alpha = 0.8) +
  scale_y_continuous(
    limits = c(0, 100),
    labels = scales::label_percent(scale = 1)
  ) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Renewables' share of electricity is rising in most countries, but falling in Ghana",
    subtitle = "Share of electricity generated from renewable sources, 2000–2025",
    x = "Year",
    y = "Renewable share of electricity",
    color = "Country",
    caption = "Source: Our World in Data, Energy Dataset (2025)"
  ) +
  theme_minimal()
```
Since 2000, the share of electricity generated from renewable sources has risen in most of the countries examined. The United Kingdom shows the fastest transition, growing from under 5% to over 50%, largely driven by the expansion of wind power, while China, Pakistan and the United States show slower but steady gains. Norway has remained close to 100% throughout because of its long-standing reliance on hydropower. Ghana is the major exception: its renewable share fell from about 90% to under 40% as rising electricity demand was met mainly by fossil fuels rather than new renewable capacity. This suggests that for fast-growing economies, the policy challenge is not only adding renewables but adding them quickly enough to keep pace with demand.










