---
title: "Assignment 4"
author: Maimoona Mohsin Subhani 
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r}
library (tidyverse)
energy <- read_csv("data/owid_energy.csv") 

```
Focusing on renewable energy in seven countries from 2000 onward 
```{r}
energy_sub <-energy |>
    filter(country %in% c("United States", "China", "India", "United Kingdom", "Norway", "Pakistan", "Ghana"),
         year >= 2000)
```

### Graph 1: Renewable Share of Electricity

```{r}
energy_sub |>
  ggplot(aes(x = year, y = renewables_electricity_share_pct, color = country)) +
  geom_line(linewidth = 1, alpha = 0.8) +
  scale_y_continuous(
    limits = c(0, 100),
    labels = scales::label_percent(scale = 1)
  ) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Renewables' share of electricity is rising in most countries, but falling in Ghana",
    subtitle = "Share of electricity generated from renewable sources, 2000–2025",
    x = "Year",
    y = "Renewable share of electricity",
    color = "Country",
    caption = "Source: Our World in Data, Energy Dataset (2025)"
  ) +
  theme_minimal()
```
Since 2000, the share of electricity generated from renewable sources has risen in most of the countries examined. The United Kingdom shows the fastest transition, growing from under 5% to over 50%, largely driven by the expansion of wind power, while China, Pakistan and the United States show slower but steady gains. Norway has remained close to 100% throughout because of its long-standing reliance on hydropower. Ghana is the major exception: its renewable share fell from about 90% to under 40% as rising electricity demand was met mainly by fossil fuels rather than new renewable capacity. This suggests that for fast-growing economies, the policy challenge is not only adding renewables but adding them quickly enough to keep pace with demand.

### Graph 2:Where Countries Stand in 2024

```{r}

energy_sub |>
  filter(year == 2024) |>
  ggplot(aes(x = renewables_electricity_share_pct,
             y = fct_reorder(country, renewables_electricity_share_pct))) +
  geom_col(fill = "#1b9e77", width = 0.7) +
  geom_text(aes(label = paste0(round(renewables_electricity_share_pct), "%")),
            hjust = -0.2, size = 3.5) +
  scale_x_continuous(limits = c(0, 110),
                     labels = scales::label_percent(scale = 1)) +
  labs(
    title = "Norway gets nearly all its electricity from renewables; India gets about a fifth",
    subtitle = "Share of electricity generated from renewable sources, 2024",
    x = "Renewable share of electricity",
    y = NULL,
    caption = "Source: Our World in Data, Energy Dataset (2025)"
  ) +
  theme_minimal()

```
In 2024, the seven countries varied widely in how much of their electricity came from renewable sources. Norway leads at 99%, and the United Kingdom is the only other country above half, at 51%. Pakistan, Ghana and China are closely grouped between 34% and 37%, despite having very different economies and energy systems. Notably, the United States ranks second to last at 24%, ahead of only India at 20%, showing that national wealth alone does not guarantee a cleaner electricity mix.Read alongside the first graph, this also shows that similar shares can hide opposite trends: Ghana and China are nearly level in 2024, but Ghana's share has fallen sharply since 2000 while China's has roughly double












